---
title: "visualization"
subtitle: "Part 2"
output: github_document
---

```{r setup}
library(tidyverse)
library(patchwork)

```

## Load the weather data
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

weather_df
```


 ## Remember this plot ..?
 
 
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5)
```
 
 ## Labels
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = 'Temperature plot 1', 
    x = 'Minimum daily temperature(C)',
    y = 'Maximum daily tempetature(C)',
    caption = 'Data from rnoaa package, temperature in 2017.'
  )
```


## Scales

Start with the same plot, x and y scales
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = 'Temperature plot 1', 
    x = 'Minimum daily temperature(C)',
    y = 'Maximum daily tempetature(C)',
    caption = 'Data from rnoaa package, temperature in 2017.'
  ) +
  scale_x_continuous(
    breaks = c(-15, 0, 15), 
    labels = c('-15 C', '0', '15')
  ) +
  scale_y_continuous(
    #trans = 'log'
    position = 'right'
  )

```

Look at color scales
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = 'Temperature plot 1', 
    x = 'Minimum daily temperature(C)',
    y = 'Maximum daily tempetature(C)',
    caption = 'Data from rnoaa package, temperature in 2017.'
  ) +
#   scale_color_hue(
#   name = 'Location',
#   h = c(100, 300))
   viridis::scale_color_viridis(
     name = 'Location',
     discrete = TRUE
   )
```

## Themes
Shift the legend.

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = 'Temperature plot 1', 
    x = 'Minimum daily temperature(C)',
    y = 'Maximum daily tempetature(C)',
    caption = 'Data from rnoaa package, temperature in 2017.'
  ) +
#   scale_color_hue(
#   name = 'Location',
#   h = c(100, 300))
   viridis::scale_color_viridis(
     name = 'Location',
     discrete = TRUE) +
  theme(legend.position = 'bottom')
```



Change the overall theme.

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = 'Temperature plot 1', 
    x = 'Minimum daily temperature(C)',
    y = 'Maximum daily tempetature(C)',
    caption = 'Data from rnoaa package, temperature in 2017.'
  ) +
#   scale_color_hue(
#   name = 'Location',
#   h = c(100, 300))
   viridis::scale_color_viridis(
     name = 'Location',
     discrete = TRUE) +
#  theme_bw()
   theme_minimal() + ##this will eliminate all things themes before
   theme(legend.position = 'bottom')
#  ggthemes::theme_economist_pal()
```


##Setting options
```{r, eval=FALSE}
library(tidyverse)

knit::opts_chunk&set(
  fig.width = 6,
  fig.asp = .6, 
  out.width = '90%'
)

theme_set(theme_minimal() + theme(legend.position = 'bottom'))  #apply this theme setting everywhere in this document

options(
  ggplot2.continuous.color = 'viridis',
  ggplot2.continuous.fill = 'viridis'
)

scale_color_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Data args in 'geom'
```{r}

central_park = 
  weather_df %>% 
  filter(name == 'CentralPark_NY')

waikiki = 
  weather_df %>% 
  filter(name == 'Waikiki_HA')

ggplot(data = waikiki, aes(x = date, y = tmax, color = name)) +
  geom_point() +
  geom_line(data = central_park)
```


##'patch work`

remenber faceting?


```{r}
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) +
  geom_density(alpha = .5) +
  facet_grid(.~name)
```

what happens when you want multipannel plots but can't facet ...?

```{r}
tmax_tmin_p = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) + 
  theme(legend.position = 'none')

prcp_dens_p = 
  weather_df %>% 
  filter(prcp > 0) %>% 
  ggplot(aes(x = prcp, fill = name)) +
  geom_density(alpha = .5) + 
  theme(legend.position = 'none')

tmax_date_p = 
  weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  theme(legend.position = 'none')

##get all these plots together
tmax_tmin_p + prcp_dens_p + tmax_date_p 
tmax_tmin_p + (prcp_dens_p + tmax_date_p)
tmax_tmin_p / (prcp_dens_p + tmax_date_p) 
(tmax_tmin_p + prcp_dens_p) / tmax_date_p 
```

## Data manipulation

Control your factors

```{r}
weather_df %>% 
  mutate(
    name = factor(name),
    name = forcats::fct_relevel(name, c('Waikiki_HA'))
  ) %>% 
  ggplot(aes(x = name, y = tmax, fill = name)) +
  geom_violin(alpha = .5)
```

What if I wanted densities for tmin and tmax simultaneously?

```{r}
weather_df %>% 
 
  pivot_longer(
    tmax:tmin,
    names_to = 'observation', 
    values_to = 'temperatures'
  ) %>% 
  ggplot(aes(x = temperatures, fill = observation)) +
  geom_density(alpha = .5) +
  facet_grid(.~name)

```

## Revisit the pupss

Data from the FAS study.

```{r}
pup_data = 
  read_csv('./data/FAS_pups.csv') %>% 
  janitor::clean_names() %>% 
  mutate(sex = recode(sex, `1` = 'male', `2` = 'female'))

litters_data = 
  read_csv('./data/FAS_litters.csv') %>% 
  janitor::clean_names() %>% 
  separate(group, into = c('dose', 'day_of_tx'), sep = 3)

fas_data = left_join(pup_data, litters_data, by = 'litter_number')



fas_data %>% 
  select(dose, day_of_tx, starts_with('pd_')) %>% 
  pivot_longer(
    pd_ears:pd_walk,
    names_to = 'outcome', 
    values_to = 'pn_day'
  ) %>% 
  drop_na() %>% 
  mutate(outcome = forcats::fct_relevel(outcome, 'pd_ears', 'pd_pivot', 'pd_walk', 'pd_eyes' )) %>% 
  ggplot(aes(x = dose, y = pn_day)) +
  geom_violin() +
  facet_grid(day_of_tx ~ outcome) 
```




